\thetitle

### 2020-11-13 sun_phases/spin_qudit_tomo.tex
added some minor discussions to appendices, renamed the classical / quantum error scales

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V \p{Q_k^V}^\dag,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k \p{Q^V_k}^\dag,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and
\begin{align}
  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
\end{align}
is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
Note that $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute $\S_V$ with $\abs{V}=2d-1$ randomly chosen measurement axes, averaged over 100 calculations of $\S_V$ or 5 minutes of runtime (for each $d$), whichever comes first.
    Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include the $O(d^3)$ runtime for pre-computing $\pi/2$-pulses $e^{\i\p{\pi/2}S_\y}$ (see Appendix \ref{sec:compute}), which can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it shot noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
This reconstruction error can be maximized analytically, yielding a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it error scale} $\epsilon_V^2$ is the same as that of computing $\S_V^2$.
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.

@@

As a final comment, we note that Eq.~\eqref{eq:bound} only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its (Euclidean) distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

We conjecture, and numerically verified for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit.
Even so, $2d-1$ measurement axes may not be the optimal number to use.
Increasing the number of measurement axes generally decreases the error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Exact reconstruction error}

@@

  = \f1n \sum_\ell \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.

@@

@@

While the error in Eq.~\eqref{eq:error_SM} cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error_SM} with the replacement $\rho\to\tilde\rho_V$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Efficiently computing error bounds}
\label{sec:compute}

[\red{todo: write a short tutorial on computing the error scale}]

>>>>

A set of measurement axes $V=\set{\v v}$ nominally induces a set of projectors $\set{\Pi_{\v v\mu}}$ that will be measured in an experiment.
By a simple change of basis, measuring these projectors is equivalent to measuring the set of transition operators $\v T_V\equiv\set{T_{\v v\ell,0}}$.
By flattening the operators in $\v T_V$ into column vectors $\oket{T_{\v v\ell,0}}$, we can construct the {\it measurement matrix}
\begin{align}
  M_V \equiv \sum_{\v v,\ell} \oket{T_{\v v\ell,0}} \bra{\v v\ell},
  \label{eq:meas_mat}
\end{align}
A necessary and sufficient condition for $V$ to allow for full state tomography is that the measured transition operators in $\v T_V$, or equivalently the columns of $M_V$, span the entire ($d^2$-dimensional) space of operators on a $d$-level spin qudit.
In this case $M_V$ must be full rank, with $d^2$ nonzero singular values.
Indexing these singular values $M^V_k$ and the corresponding (normalized) right singular vectors $\v x^V_k\equiv(x^V_{k,1},x^V_{k,2},\cdots,x^V_{k,\abs{V}})$ by an integer $k\in\set{1,2,\cdots,d^2}$, we can construct the orthonormal qudit operators
\begin{align}
  Q^V_k \equiv \sum_j q^V_{kj} T_j,
  &&
  \v q^V_k \equiv \f{\v x^V_k}{M^V_k},
\end{align}
where for shorthand we use a combined index $j=\p{\v v_j,\ell_j}$ to specify both a measurement axis $\v v_j$ and a degree $\ell_j$, which identify the transition operator $T_j\equiv T_{\v v_j\ell_j,0}$.
These operators allow us to expand any state $\rho$ of a $d$-level spin qudit in the form
\begin{align}
  \rho = \sum_{k=1}^{d^2} \rho_k^V \p{Q_k^V}^\dag,
  &&
  \rho_k^V \equiv \bk{Q^V_k}_\rho.
\end{align}
Given empirical estimates $\tilde T_j$ of the expectation values $\bk{T_j}_\rho$, an empirical estimate $\tilde\rho_V$ of $\rho$ is then
\begin{align}
  \tilde\rho_V \equiv \sum_k \tilde\rho^V_k \p{Q^V_k}^\dag,
  \label{eq:reconstructed_state}
\end{align}
where
\begin{align}
  \tilde\rho^V_k \equiv \sum_j q^V_{kj} \tilde T_j
  \approx \sum_j q^V_{kj} \bk{T_j}_\rho
  = \bk{Q^V_k}_\rho
  = \rho^V_k.
\end{align}
The measurement matrix $M_V$ allows us to make concrete statements about the statistical error between the empirical estimate $\tilde\rho_V$ and the true state $\rho$.
Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance at most $\epsilon^2$:
\begin{align}
  \tilde T_j = \bk{T_j}_\rho + \epsilon_j,
  &&
  \bbk{\epsilon_j \epsilon_{j'}} \le \epsilon^2 \delta_{jj'},
  \label{eq:error_assumption}
\end{align}
where $\set{\epsilon_j}$ are independent random variables and we use the double brackets $\bbk{\cdot}$ to denote statistical averaging.
In this case, the mean squared error with which $\tilde\rho^V_k$ approximates $\rho^V_k$ is
\begin{align}
  \Bbk{\abs{\tilde\rho^V_k-\rho^V_k}^2}
  &= \Bbk{\p{\tilde\rho^V_k-\rho^V_k}^*\p{\tilde\rho^V_k-\rho^V_k}} \\
  &= \sum_{j,j'} \p{q^V_{kj}}^* q^V_{kj'}\,
  \bbk{\epsilon_j \epsilon_{j'}} \\
  &\le \sum_j \abs{q^V_{kj}}^2 \epsilon^2
  = \p{\f{\epsilon}{M^V_k}}^2.
\end{align}
Using the fact that the operators $Q^V_k$ are orthonormal, we can therefore bound the mean squared (Euclidean) distance between $\tilde\rho_V$ and $\rho$ as
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  \le \epsilon^2 \S_V^2,
  \label{eq:bound_eps}
\end{align}
where $\norm{X} \equiv \sqrt{\obk{X|X}} = \sqrt{\tr\p{X^\dag X}}$ is the (Hilbert-Schmidt, or Frobenius) norm of $X$, and the {\it classical error scale}
\begin{align}
  \S_V^2 \equiv \sum_k\f1{\p{M^V_k}^2} = \norm{M_V^{-1}}^2,
\end{align}
is the squared norm of the right inverse $M_V^{-1}$ of $M_V$.
Note that the classical error scale $\S_V^2$ diverges if $M_V$ is singular, in which case measuring spin projections along all axes in $V$ does not provide sufficient information to reconstruct arbitrary quantum states.

Computing the classical error scale $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
The complexity of this task can be greatly reduced by the fact that the degree $\ell$ of a transition operator $T_{\ell m}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for details on efficiently computing $\S_{V\ell}^2$).
If we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure \ref{fig:times}).

\begin{figure}
  \centering
  \includegraphics{error_scale/error_times.pdf}
  \caption{Serial runtime to compute $\S_V$ with $\abs{V}=2d-1$ randomly chosen measurement axes, averaged over 100 calculations of $\S_V$ or 5 minutes of runtime (for each $d$), whichever comes first.
    Dashed line shows a fit to runtime $t=c d^\alpha$ for all $d\ge100$, finding $\alpha\approx 3.7$.
    These results do not include fixed $O(d^3)$ runtimes for pre-computing quantities that can be recycled for every new choice of measurement axes $V$.}
  \label{fig:times}
\end{figure}

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} in Appendix \ref{sec:error}, where we instead assume that $\tilde\rho_V$ is built from $n$ independent measurements of spin projection along every axis $\v v\in V$, with shot noise the dominant source of error.
In this case, the mean squared distance between the approximation $\tilde\rho_V$ and the true state $\rho$ can be compactly expressed as as an inner product of the {\it covariance matrices}
\begin{align}
  \C_\ell\sp{\rho} \equiv \sum_{m,n}
  \cov_\rho\p{T_{\ell m}^\dag, T_{\ell n}} \op{m}{n}
\end{align}
where $\cov_\rho\p{X,Y} \equiv \bk{XY}_\rho - \bk{X}_\rho \bk{Y}_\rho$, and the {\it noise matrices}
\begin{align}
  \N_{V\ell} \equiv M_{V\ell} \diag\sp{M_{V\ell}^{-1}
    \p{M_{V\ell}^{-1}}^\dag} M_{V\ell}^\dag,
\end{align}
where $\diag\sp{X}$ sets all off-diagonal parts of $X$ to zero:
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2}
  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.
  \label{eq:error}
\end{align}
This reconstruction error can be maximized analytically, yielding a reconstruction error bound of the form
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} \le \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the complexity of computing the {\it quantum error scale} $\epsilon_V^2$ is the same as that of computing the classical error scale $\S_V^2$.
For comparison with the bound in Eq.~\eqref{eq:bound_eps}, we note that the error scale $\epsilon_V^2<\S_V^2/2$, so the previous bound still holds with the replacement $\epsilon^2\to1/2n$.

@@

As a final comment, we note that the quantum error scale only provides a pessimistic upper bound on statistical error, which can be calculated without prior knowledge of the true qudit state $\rho$.
The actual error in the reconstruction $\tilde\rho_V$ of a particular state $\rho$, provided by Eq.~\eqref{eq:error}, may be considerably smaller.
While the true error in $\tilde\rho_V$ cannot be known exactly without knowing $\rho$, this error can be estimated by computing the right-hand side of Eq.~\eqref{eq:error} with the replacement $\rho\to\tilde\rho_V$.

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following simple protocol for spin qudit tomography: select a random set of measurement axes $V$ by uniformly sampling points on the sphere \footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and optimize (using any standard minimization algorithm) the $2\abs{V}$ parameters in $V$ (two angles for each point $\v v\in V$) to minimize the quantum error scale $\epsilon_V^2$ defined in Eq.~\eqref{eq:scale}.
If $\abs{V}$ is too large for such optimization  to be practical, one can instead generate many random sets of measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V^2$.
After choosing a set of measurement axes $V$, (experimentally) make $n$ measurements of spin projection along every axis $\v v\in V$, and set $\tilde\Pi_{\v v\mu}\approx\bk{\Pi_{\v v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
The estimates $\tilde\Pi_{\v v\mu}$ of $\bk{\Pi_{\v v\mu}}_\rho$ determine estimates of the expectation values $\bk{T_{\v v\ell,0}}_\rho$, which can be used to construct the state $\tilde\rho_V$ defined in Eq.~\eqref{eq:reconstructed_state}.
If $\tilde\rho_V$ has negative eigenvalues, its (Euclidean) distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}.

@@

As a last point, we mention that the quantum error scale $\epsilon_V^2$ can also be used to evaluate the ``quality'' of measurement axes $V$ chosen with existing methods for sparse signal recovery on the sphere \cite{rauhut2011sparse, alem2012sparse}.
In practice, we find that uniform random sampling outperforms these methods, in the sense that it generates low-error-scale sets of measurement axes with higher probability.
Two observations may explain the poor performance of these sparse signal recovery methods.
First, spin qudit tomography essentially boils down to recovering sparse signals with a particular harmonic structure, namely a fixed degree $\ell$.
The existing sparse signal recovery algorithms do not assume or account for any harmonic structure, and may therefore perform worse than expected on this particular subset of signals.
Second, spin qudit tomography requires recovering signals that are band-limited at degree $d$, but have $O(d)$ non-zero harmonic coefficients (out of $d^2$ coefficients in total).
The existing sparse signal recovery methods, on the other hand, may only perform well for signals with sparsity much smaller than $d$.
In any case, the fact that uniform sampling outperforms these methods only simplifies our randomized tomography protocol.

@@

We conjecture, and numerically verified for $d\le200$, that measurements of spin projection along $2d-1$ axes suffice to reconstruct an arbitrary state of a $d$-level spin qudit.
Even so, $2d-1$ measurement axes may not be the optimal number to use.
Increasing the number of measurement axes generally decreases the quantum error scale $\epsilon_V^2$, but at the cost of having to estimate more observables.
At a fixed total number of measurements, increasing the number of measurement axes $\abs{V}$ reduces the number of measurements $n$ devoted to each axis $\v v\in V$.
This trade-off begs the question: how many measurement axes should be chosen to perform spin qudit tomography?

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Exact reconstruction error}
\label{sec:error_exact}

@@

  = \f1n \sum_{\ell>0} \obk{\N_{V\ell}|\C_\ell\sp{\rho}}.

@@

Here we explicitly throw out the $\ell=0$ term because $\C_0\sp{\rho}=0$.

@@

We discuss the details of efficiently computing the error in Eq.~\eqref{eq:error_SM} in Appendix \ref{sec:compute}, but devote the rest of this Appendix to finding upper bounds for this error.

@@

The naive error bound in Eq.~\eqref{eq:bound_eps} of the main text therefore holds in the quantum case with the replacement $\epsilon^2\to1/2n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Efficiently computing reconstruction errors and bounds}
\label{sec:compute}

[\red{todo: write a short tutorial on computing error bounds}]

% We note that when computing this bound in practice, the vector $\oket{\I g_{L\ell}}$ should be pre-computed once and recycled for every new choice of axes $V$, and the inner product $\obk{\N_{V\ell}|\D_M|\I g_{L\ell}}$ to compute $\chi^V_{LM}$ should be seen as an inner product of $O(d)$-sized off-diagonal bands of $\I g_{L\ell}$ and $\N_{V\ell}$.
% Also, the construction of $\oket{\chi_V}$, $\L_{V\ell}^{-1}$, and $\oket{\rho_V^\star}$ is most efficiently performed in the transition operator basis $\set{\oket{T_{\ell m}}}$.
